Incentivizing Exploration by Heterogeneous Users

### 2018-02-09 algorithm.tex
edited the payment, regret and lower bound proof

\begin{itemize}
\item $i$ has been pulled at most%
\footnote{in fact: exactly, since the algorithm entered phase $s$}
$s-1$ times up to time $t$, i.e., $\NumPull{t}{s} \leq s-1$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms'
attribute vectors, the probability of pulling arm $i$ is less than $1/s$.
Let $\PullProb{t}{i} = \Prob[\AgV \sim \AgentDist]{\AgV \cdot \ArmEV{t}{i} > \AgV
  \cdot \ArmEV{t}{i'} \mbox{ for all } i' \neq i}$
be the probability that arm $i$ will be pulled
by the next (random) agent based on the current estimates. 
Then, the second criterion for payment-eligibility is that
$\PullProb{t}{i} < 1/s$.
\end{itemize}

When multiple arms are payment-eligible, the algorithm picks one
uniformly at random to incentivize.
When the algorithm decides to incentivize an arm $i$,
it offers ``whatever it takes,'' i.e., offers a payment of
$\Pay{t}{i} = \max_{\AgV,i'} \AgV \cdot (\ArmEV{t}{i'} - \ArmEV{t}{i})$.
The maximum for \AgV is taken over the support of \AgentDist;
recall that we assumed this support to be compact.
The payment \Pay{t}{i} may appear unnecessarily high.
Indeed, we show in Section~\ref{sec:pi} that it suffices to
incentivize only a $1/s$ fraction of the agents.
\dkcomment{Given that we don't prove anything there, is ``show'' the
  right word?}
Here, we focus on the higher payments for simplicity of analysis.

@@

\begin{proof}
Let $S$ be the set of arms that have been pulled at most
$s-1$ times at the start of phase $s$.
Thus, phase $s$ ends when each $i \in S$ has been pulled at least once.
Consider any time step $t$.
If $S$ contains at least one payment-eligible arm $i$,
then some payment-eligible arm will be pulled with probability 1,
and every payment-eligible arm is in $S$.
In particular, the size of $S$ decreases with probability 1.
If $S$ contains no payment-eligible arm,
then by definition of payment-eligibility,
each arm $i \in S$ has probability at least $1/s$ of being pulled.
In particular, the size of $S$ decreases with probability at least $1/s$.

In either case, the expected number of time steps until
$|S|$ decreases is at most $s$,
so the expected number of steps until $S$ is empty is at most
$\ARMNUM \cdot s$.
\dkcomment{If we wanted, this bound could probably be improved using a
  coupon collector argument. But it would look uglier.}
\end{proof}

@@

\begin{proof}
We partition phases into early and late phases.
For each of the early phases,
we crudely bound the number of payments by \ARMNUM,
using that each arm is incentivized at most once per phase.
For later phases,
we use that arms are payment-eligible only if their myopic pulling
probability is less than $1/s$.
Since at least a \MinProb fraction of agents would prefer arm $i$ if
given accurate estimates, this would require misestimates that make a
significant fraction of agents pull the wrong arm.
But in later phases, such misestimates are unlikely by
Corollary~\ref{cor:right-choice}.

To make this intuition precise, we set
$\delta = \frac{\MinProb}{2 \TieDensity}$,
and $x = \frac{\delta}{2 \Diam d}$.
The cutoff point between early and late phases is now set to
$\EvenLaterPhase = \max(50, \frac{92.16 \sigma^4}{x^4}, \frac{2}{\MinProb})$.
Consider the \Kth{s} pull of an arm $i$, for $s \geq \EvenLaterPhase$.
If this pull occurs before phase $s$, it was definitely not incentivized.
So we condition on the pull occurring in phase $s$,
and let $\tau$ be the time step when this pull occurred; notice that
this means that this must be the first pull of arm $i$ in phase $s$.
We can now apply Corollary~\ref{cor:right-choice}
with this choice of $x$,
and obtain that with probability at least 
$1 - 24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s x^2}{16 \sigma^2} \right)$,
all agent types \AgV with
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}})
> 2\Diam d x = \delta$
will pull their best arm \Best{\AgV}.
\dkcomment{We are losing a factor \ARMNUM here.
We already took a union bound over all arms in the lemma, and
we now apply it only to one arm, then sum over all arms.}

@@

Thus, for any phase $s \geq \EvenLaterPhase$
Algorithm~\ref{alg:basic-incentivizing} incentivizes any given arm $i$
with probability at most
$24 \ARMNUM d \cdot \exp \left(
  -\frac{1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
\right)$.
Summing over all arms and phases $s$,
and adding the at most $\ARMNUM \EvenLaterPhase$ incentivizations in
the first \EvenLaterPhase phases, 
the expected total number of arm incentivizations is at most
\begin{align*}
\ARMNUM \EvenLaterPhase
  + 24 \ARMNUM^2 d \cdot \frac{1}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)}
& = O\left( \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right).
\end{align*}
\dkcomment{It seems like with our choice of $x$ (which may be
  unavoidable), we get quite large factors, in particular with the
  fourth power.
  Also, it would be nice if we could simplify the $\frac{1}{1-\exp()}$
  term. Unfortunately, I think that the inequality goes the wrong way
  for applying the standard $\exp(-x) \geq 1-x$ bound.}

@@

\begin{proof}
Regret can arise in two ways:
(1) an agent was incentivized to pull a suboptimal arm, or
(2) an agent myopically pulled a suboptimal arm.
By Lemma~\ref{lem:numP}, Algorithm~\ref{alg:basic-incentivizing}
incentivizes agents at most 
$O \left( \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$
times in expectation, each time causing regret at most \MAXR,
for a total expected regret of at most
$O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$.
For the rest of the proof, we focus on the regret incurred when agents
pull arms myopically and make mistakes.

@@

We first consider agents \AgV incurring positive regret less than $g(s)$.
The second-best arm \Second{\AgV} of \AgV must satisfy
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq g(s)$.
By Lemma~\ref{lem:sdelta}, the total measure of such agents
is $\AlmostTied{g(s)} \leq \TieDensity \cdot g(s)$,
and each incurs regret at most $g(s)$.
Summing over all time steps, the expected total regret from such agents is at most
\begin{align}
&E[\sum_{s=1}^{T}\sum_{t=t_{s-1}+1}^{t_{s}} \TieDensity \cdot g^2(s)] \leq \sum_{s=1}^{T}\TieDensity\cdot g^2(s)Ns. \nonumber 
\end{align}

@@

\begin{align*}
&O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)
\\
+ & \sum_{s = 1}^T 
  \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR
  \cdot \ARMNUM s,
+ \sum_{s=1}^{T}\TieDensity \cdot g^2(s)\cdot Ns.
\end{align*}

@@

\begin{align*}
\lefteqn{\sum_{s=1}^{T} 24\ARMNUM^2 d s \MAXR \exp\left(-2\log(T)\right)  +
          \frac{128\Diam^2 d^2\sigma^2
          \TieDensity\log(T)}{1.8}\sum_{s=1}^{T}\frac{1}{s}}
\\ & \leq
24\ARMNUM^2 d R\frac{T(T-1)}{2T^2}  + 71.12 \Diam^2 d^2\sigma^2
          \TieDensity\log(T)(\log(T)+1)
  \\ & \leq
12 \ARMNUM^2 d R  + 71.12 \Diam^2 d^2\sigma^2 \TieDensity\log(T)(\log(T)+1). 
\end{align*}

Adding all terms, and substituting the definition of
$\LatePhase = O(\frac{\sigma^4}{x^4})$
now gives an upper bound of

\begin{multline*}
O(\ARMNUM \MAXR \LatePhase^2) + 
O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right) +
O(\Diam^2 d^2\sigma^2 \TieDensity\log(T)(\log(T)+1)) 
\end{multline*}

@@

\begin{proof}
The proof of this corollary is similar to the proof of Theorem~\ref{rst:regret}. If we set $g^2(t)=\frac{2\log(T)\times 64\Diam^2 d^2\sigma^2}{1.8t}$, then there exists a $t^{0}$ such that for $t>t_{0}$, $q(g(t))=0$. Thus, similar to equation~\eqref{equ:regret}, we know the cumulative expected regret when we let the agents play myopically is bounded above by
\begin{align}
\sum_{s=1}^{\LatePhase}\ARMNUM \MAXR s + \sum_{s=1}^{T} 24\ARMNUM d\exp\left(-\frac{1.8s f(s)^2}{64 \Diam^2 d^2\sigma^2}\right)\times \MAXR \times \ARMNUM s+ \sum_{t=1}^{t^{0}}\TieDensity g(t)^2 \nonumber
\end{align}
Therefore, based on the same analysis of Theorem~\ref{rst:regret}, we know the cumulative regret is bounded by $O(\ARMNUM^2)$.
\end{proof}

>>>>

\begin{itemize}
\item $i$ has been pulled at most%
\footnote{in fact: exactly, since the algorithm entered phase $s$}
$s-1$ times up to time $t$, i.e., $\NumPull{t}{s} \leq s-1$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms'
attribute vectors, the probability of pulling arm $i$ is less than $1/\log(s)$. \bccomment{Do we need to mention what if $\frac{1}{\log(s)}>1$? }
Let $\PullProb{t}{i} = \Prob[\AgV \sim \AgentDist]{\AgV \cdot \ArmEV{t}{i} > \AgV
  \cdot \ArmEV{t}{i'} \mbox{ for all } i' \neq i}$
be the probability that arm $i$ will be pulled
by the next (random) agent based on the current estimates. 
Then, the second criterion for payment-eligibility is that
$\PullProb{t}{i} < 1/\log(s)$.
\end{itemize}

When multiple arms are payment-eligible, the algorithm picks one
uniformly at random to incentivize.
When the algorithm decides to incentivize an arm $i$,
it offers ``whatever it takes,'' i.e., offers a payment of
$\Pay{t}{i} = \max_{\AgV,i'} \AgV \cdot (\ArmEV{t}{i'} - \ArmEV{t}{i})$.
The maximum for \AgV is taken over the support of \AgentDist;
recall that we assumed this support to be compact.
The payment \Pay{t}{i} may appear unnecessarily high.
Indeed, we show in Section~\ref{sec:pi} that it suffices to
incentivize only a $1/\log(s)$ fraction of the agents.
\dkcomment{Given that we don't prove anything there, is ``show'' the
  right word?}
Here, we focus on the higher payments for simplicity of analysis.

@@

\begin{proof}
Let $S$ be the set of arms that have been pulled at most
$s-1$ times at the start of phase $s$.
Thus, phase $s$ ends when each $i \in S$ has been pulled at least once.
Consider any time step $t$.
If $S$ contains at least one payment-eligible arm $i$,
then some payment-eligible arm will be pulled with probability 1,
and every payment-eligible arm is in $S$.
In particular, the size of $S$ decreases with probability 1.
If $S$ contains no payment-eligible arm,
then by definition of payment-eligibility,
each arm $i \in S$ has probability at least $1/\log(s)$ of being pulled.
In particular, the size of $S$ decreases with probability at least $1/\log(s)$.

In either case, the expected number of time steps until
$|S|$ decreases is at most $s$,
so the expected number of steps until $S$ is empty is at most
$\ARMNUM \cdot \log(s)$.
\dkcomment{If we wanted, this bound could probably be improved using a
  coupon collector argument. But it would look uglier.}
\end{proof}

@@

\begin{proof}
We partition phases into early and late phases.
For each of the early phases,
we crudely bound the number of payments by \ARMNUM,
using that each arm is incentivized at most once per phase.
For later phases,
we use that arms are payment-eligible only if their myopic pulling
probability is less than $1/\log(s)$.
Since at least a \MinProb fraction of agents would prefer arm $i$ if
given accurate estimates, this would require misestimates that make a
significant fraction of agents pull the wrong arm.
But in later phases, such misestimates are unlikely by
Corollary~\ref{cor:right-choice}.

To make this intuition precise, we set
$\delta = \frac{\MinProb}{2 \TieDensity}$,
and $x = \frac{\delta}{2 \Diam d}$.
The cutoff point between early and late phases is now set to
$\EvenLaterPhase = \max(50, \frac{92.16 \sigma^4}{x^4}, \exp(\frac{2}{\MinProb}))$.
Consider the \Kth{s} pull of an arm $i$, for $s \geq \EvenLaterPhase$.
If this pull occurs before phase $s$, it was definitely not incentivized.
So we condition on the pull occurring in phase $s$,
and let $\tau$ be the time step when this pull occurred; notice that
this means that this must be the first pull of arm $i$ in phase $s$.
We can now apply Corollary~\ref{cor:right-choice}
with this choice of $x$,
and obtain that with probability at least 
$1 - 24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s x^2}{16 \sigma^2} \right)$,
all agent types \AgV with
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}})
> 2\Diam d x = \delta$
will pull their best arm \Best{\AgV}.
\dkcomment{We are losing a factor \ARMNUM here.
We already took a union bound over all arms in the lemma, and
we now apply it only to one arm, then sum over all arms.}

@@

Thus, for any phase $s \geq \EvenLaterPhase$
Algorithm~\ref{alg:basic-incentivizing} incentivizes any given arm $i$
with probability at most
$24 \ARMNUM d \cdot \exp \left(
  -\frac{1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
\right)$.
Summing over all arms and phases $s$,
and adding the at most $\ARMNUM \EvenLaterPhase$ incentivizations in
the first \EvenLaterPhase phases, 
the expected total number of arm incentivizations is at most
\begin{align*}
\ARMNUM \EvenLaterPhase
  + 24 \ARMNUM^2 d \cdot \frac{1}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)}
& = O\left( \max\{\frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}, N\exp(\frac{2}{p})\}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right).
\end{align*}
\dkcomment{It seems like with our choice of $x$ (which may be
  unavoidable), we get quite large factors, in particular with the
  fourth power.
  Also, it would be nice if we could simplify the $\frac{1}{1-\exp()}$
  term. Unfortunately, I think that the inequality goes the wrong way
  for applying the standard $\exp(-x) \geq 1-x$ bound.}

@@

\begin{proof}
Regret can arise in two ways:
(1) an agent was incentivized to pull a suboptimal arm, or
(2) an agent myopically pulled a suboptimal arm.
By Lemma~\ref{lem:numP}, Algorithm~\ref{alg:basic-incentivizing}
incentivizes agents at most 
$O \left( \max\{\frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}, N\exp(\frac{2}{p})\}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$
times in expectation, each time causing regret at most \MAXR,
for a total expected regret of at most
$O \left(\MAXR\cdot\max\{\frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}, N\exp(\frac{2}{p})\}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$.
For the rest of the proof, we focus on the regret incurred when agents
pull arms myopically and make mistakes.

@@

We first consider agents \AgV incurring positive regret less than $g(s)$.
The second-best arm \Second{\AgV} of \AgV must satisfy
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq g(s)$.
By Lemma~\ref{lem:sdelta}, the total measure of such agents
is $\AlmostTied{g(s)} \leq \TieDensity \cdot g(s)$,
and each incurs regret at most $g(s)$.
Summing over all time steps, the expected total regret from such agents is at most
\begin{align}
&E[\sum_{s=1}^{T}\sum_{t=t_{s-1}+1}^{t_{s}} \TieDensity \cdot g^2(s)] \leq \sum_{s=1}^{T}\TieDensity\cdot g^2(s)N\log(s). \nonumber 
\end{align}

@@

\begin{align}
&O \left( \MAXR\cdot \max\{\frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}, N\exp(\frac{2}{p})\}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)
\\
+ & \sum_{s = 1}^T 
  \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR
  \cdot \ARMNUM s,
+ \sum_{s=1}^{T}\TieDensity \cdot g^2(s)\cdot N\log(s). \label{thm:equ}
\end{align}

@@

\begin{align*}
\lefteqn{\sum_{s=1}^{T} 24\ARMNUM^2 d s \MAXR \exp\left(-2\log(T)\right)  +
          \frac{288\ARMNUM\Diam^2 d^2\sigma^2
          \TieDensity\log(T)}{1.8}\sum_{s=1}^{T}\frac{\log(s)}{s}}
\\ & \leq
24\ARMNUM^2 d R\frac{T(T-1)}{2T^2}  + 160 \ARMNUM \Diam^2 d^2\sigma^2
          \TieDensity(\log(T))^2(\log(T)+1)
  \\ & \leq
12 \ARMNUM^2 d R  + 160 \ARMNUM\Diam^2 d^2\sigma^2 \TieDensity(\log(T))^2(\log(T)+1). 
\end{align*}

Adding all terms gives an upper bound of

\begin{multline*}
O \left(\MAXR\cdot \max\{\frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}, N\exp(\frac{2}{p})\}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right) +
O(\ARMNUM\Diam^2 d^2\sigma^2 \TieDensity(\log(T))^3) 
\end{multline*}

@@

\begin{proof}
The proof of this corollary is similar to the proof of Theorem~\ref{rst:regret}. If we set $g^2(s) = \frac{288\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$, then there exists a $s^{0}$ such that for $s>s_{0}$, $q(g(s))=0$. Thus, similar to equation~\eqref{thm:equ}, we know the cumulative expected regret when we let the agents play myopically is bounded above by
\begin{align}
\sum_{s=1}^{T} 24\ARMNUM^2 d s \MAXR \exp\left(-2\log(T)\right)+ \sum_{s=1}^{s^{0}}\TieDensity g(s)^2\ARMNUM\log(s) \nonumber
\end{align}
Therefore, based on the same analysis of Theorem~\ref{rst:regret}, we know the cumulative regret is bounded by $O(\ARMNUM^2)$.
\end{proof}
